Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2018-08-06 submission/manuscript.Rmd
complete Ada comments on results

**Random Forest models for SCFA concentrations have different top 10 most important OTUs than tumor models.** Using OTU data we built Random Forest models to classify either SCFA concentration or higher/lower than median SCFA concentrations. Overall, OTU data had a reasonable ability to classify both SCFA concentrations and high/low SCFA groups [Figure 5A & S2A]. However, these models tended to be over fit, suggesting that rarer taxa may be important for this classification [Figure 5A & S2A]. Additionally, there were differences in accuracy for both model types based on whether the individual had no tumors, an adenoma, or a carcinoma [Figure 5B & S2B]. There also was minimal overlap between these SCFA classification model's most important OTUs and those used to classify patients with adenoma or carcinoma tumors [Figure 4B-C, 4E-F, 5C-E, and S2C-E]. The only OTU that did overlap between the models was OTU00167 (*Clostridiales*) [Figure 4B-C, 4E-F, 5C-E, and S2C-E]. Additionally, OTU00167 was in the top 10 most important OTUs for the OTU adenoma model but not in the SCFA and OTU adenoma model while acetate and butyrate concentrations were [Figure 4B-C]. These observations provide further evidence that it is possible to identify specific OTUs associated with higher SCFA concentrations and that these OTUs belong to taxa known to produce acetate, butyrate, and propionate. Although it is possible to identify OTUs associated with SCFA production, our results do not support the hypothesis that SCFA concentration or OTUs associated with their production are different between individuals with no tumors and patients with adenomas or carcinomas.

>>>>

**The most important OTUs in Random Forest models built to classify SCFA concentrations or tumors are not the same.** It is possible that due to the way the Random Forest algorithm works, OTUs associated with SCFA concentrations could be downweighted in importance within the adenoma or carcinoma models when SCFA concentrations are included. To test if this is the case we used OTU data and built Random Forest models to classify SCFA concentrations. Although the correlation between the predicted and actual SCFA concentrations were moderately associated with each other [Figure 5A], there was a difference in accuracy based on whether the fecal sample was from an individual without tumors or from patients with adenomas or carcinomas [Figure 5B]. Additionally, because the training set R^2^ was always higher than the test set R^2^, all SCFA concentration models tended to be over fit, suggesting that rarer taxa are important for these classifications [Figure 5A]. When comparing the adenoma or carcinoma model to the SCFA concentration models there was minimal overlap between these model's most important OTUs [Figure 4B-C, 4E-F and 5C-E]. The only OTU that was in the top 10 most important variables that had overlap between the models was OTU00167 (*Clostridiales*) [Figure 4B-C, 4E-F, 5C-E]. Similar observations were made when using high/low SCFA groups based on higher/lower than median SCFA concentrations [Figure S2]. Collectively, these observations provide evidence that it is possible to identify specific OTUs associated with higher SCFA concentrations and accordingly these OTUs belong to taxa known to produce acetate, butyrate, and propionate. Although it is possible to identify OTUs associated with SCFA production, overall, our results do not support the hypothesis that SCFA concentration or OTUs associated with their production are different between individuals with no tumors and patients with adenomas or carcinomas.

### 2018-08-06 submission/manuscript.tex
complete Ada comments on results

\textbf{Random Forest models for SCFA concentrations have different top
10 most important OTUs than tumor models.} Using OTU data we built
Random Forest models to classify either SCFA concentration or
higher/lower than median SCFA concentrations. Overall, OTU data had a
reasonable ability to classify both SCFA concentrations and high/low
SCFA groups {[}Figure 5A \& S2A{]}. However, these models tended to be
over fit, suggesting that rarer taxa may be important for this
classification {[}Figure 5A \& S2A{]}. Additionally, there were
differences in accuracy for both model types based on whether the
individual had no tumors, an adenoma, or a carcinoma {[}Figure 5B \&
S2B{]}. There also was minimal overlap between these SCFA classification
model's most important OTUs and those used to classify patients with
adenoma or carcinoma tumors {[}Figure 4B-C, 4E-F, 5C-E, and S2C-E{]}.
The only OTU that did overlap between the models was OTU00167
(\emph{Clostridiales}) {[}Figure 4B-C, 4E-F, 5C-E, and S2C-E{]}.
Additionally, OTU00167 was in the top 10 most important OTUs for the OTU
adenoma model but not in the SCFA and OTU adenoma model while acetate
and butyrate concentrations were {[}Figure 4B-C{]}. These observations
provide further evidence that it is possible to identify specific OTUs
associated with higher SCFA concentrations and that these OTUs belong to
taxa known to produce acetate, butyrate, and propionate. Although it is
possible to identify OTUs associated with SCFA production, our results
do not support the hypothesis that SCFA concentration or OTUs associated
with their production are different between individuals with no tumors
and patients with adenomas or carcinomas.

>>>>

\textbf{The most important OTUs in Random Forest models built to
classify SCFA concentrations or tumors are not the same.} It is possible
that due to the way the Random Forest algorithm works, OTUs associated
with SCFA concentrations could be downweighted in importance within the
adenoma or carcinoma models when SCFA concentrations are included. To
test if this is the case we used OTU data and built Random Forest models
to classify SCFA concentrations. Although the correlation between the
predicted and actual SCFA concentrations were moderately associated with
each other {[}Figure 5A{]}, there was a difference in accuracy based on
whether the fecal sample was from an individual without tumors or from
patients with adenomas or carcinomas {[}Figure 5B{]}. Additionally,
because the training set R\textsuperscript{2} was always higher than the
test set R\textsuperscript{2}, all SCFA concentration models tended to
be over fit, suggesting that rarer taxa are important for these
classifications {[}Figure 5A{]}. When comparing the adenoma or carcinoma
model to the SCFA concentration models there was minimal overlap between
these model's most important OTUs {[}Figure 4B-C, 4E-F and 5C-E{]}. The
only OTU that was in the top 10 most important variables that had
overlap between the models was OTU00167 (\emph{Clostridiales}) {[}Figure
4B-C, 4E-F, 5C-E{]}. Similar observations were made when using high/low
SCFA groups based on higher/lower than median SCFA concentrations
{[}Figure S2{]}. Collectively, these observations provide evidence that
it is possible to identify specific OTUs associated with higher SCFA
concentrations and accordingly these OTUs belong to taxa known to
produce acetate, butyrate, and propionate. Although it is possible to
identify OTUs associated with SCFA production, overall, our results do
not support the hypothesis that SCFA concentration or OTUs associated
with their production are different between individuals with no tumors
and patients with adenomas or carcinomas.
